### 2016-04-23 manuscript.tex
More changes

\section{Weighted Linkage Disequilibrium} As \cite{loh2013inferring} noted, we
cannot use the LD in the admixed population directly, because the allele
frequency differences in the source populations can be of either sign. As in
\cite{loh2013inferring} , we solve this problem by computing the product of the
values of the three-point linkage disequilibrium coefficient with the product of
the allele frequency differences. Using equation \ref{local_ancestry} we obtain $$
\delta_x\delta_y\delta_z
D_3(d,d')=\delta_x^2\delta_y^2\delta_z^2\mathbb{E}[\text{cov}(A_x,A_y,A_z)], $$
because the local ancestry in the admixed sample is independent of the allele
frequencies in the admixed population. For inference purposes, we estimate this
function by averaging over triples of SNPs which are separated by distances of
approximately $d$ and $d'$. The LD term is estimated from the admixed
population, while the $\delta$'s are estimated from reference populations which
are closely related to the two source populations. We notice that both this
approach, as well as the previous approaches (e.g., \cite{loh2013inferring} ),
does not take genetic drift in the source populations after the time of
admixture into account, i.e. there is an assumption of both this method and
previous methods that the allele frequencies in the ancestral source populations
can be approximated well using the allele frequencies in the extant populations.

We arrange the data from the admixed samples in an $n\times S_n $ matrix
$\mathbf{H}$, where $n$ is the number of admixed haplotypes/genotypes, and $S_n$
is the number of segregating sites in the sample. For ease of notation, we
assume that the positions are given in units which make the unit interval equal
to the desired bin resolution.

For a given $d$ and $d'$ the set of SNP triples we use in the estimator for the
weighted LD is $$ S[d,d'] = \left\{x,y,z: d\leq  x-y < d+1 \text{ and } d' \leq
y-z < d'+1 \right\}. $$ Let $f_x$ be empirical allele frequency in the admixed
population. An unbiased estimator of the weighted three-point linkage
disequilibrium coefficient is then \begin{align*} \hat{a}[d,d']&=
\frac{1}{|S[d,d']|} \sum_{x,y,z\in S[d,d']}\frac{n\sum_{i=1}^n
\delta_x\delta_y\delta_z(H_{i,x}-f_x)(H_{i,y}-f_y)(H_{i,z}-f_z)}{(n-1)(n-2)}.
\end{align*}

@@

First, rearrange $\hat{a}$ to get
\begin{align*}
	\hat{a}[d,d'] &=
	\frac{n}{(n-1)(n-2)}
	\frac{
		\sum_{i=1}^n \sum_{x,y,z\in S[d,d']}
		\delta_x \delta_y \delta_z (H_{i,x} - f_x) (H_{i,y} - f_y) (H_{i,z} - f_z)
	}{
		\sum_{x,y,z \in S[d,d']}
		1
	},
\end{align*}
and define sequences $b_i[d]$ and $c[d]$ by binning the data and then doubling
the length by padding with $P$ zeros,
\begin{align*}
	b_i[d] &=
	\left\{
		\begin{array}{ll}
			\sum_{x: d\leq \lfloor{x}\rfloor < d+1} \delta_x(H_{i,x}-f_x) & : 0 \leq d < P\\
			0 & : P\leq d <2P
		\end{array}
	\right.\\
	c[d] &=
	\left\{
		\begin{array}{ll}
			|\{x: d\leq \lfloor{x}\rfloor < d+1\}| & : 0\leq d < P\\
			0 & : P\leq d <2P
		\end{array}
	\right.
\end{align*}
We can approximate $|S[d,d']|$ and the $n$ sums in the numerator of
$\hat{a}[d,d']$ in terms of convolutions of these sequences:
$$
	|S[d,d']| \approx \sum_{w=0}^P c[w]c[w+d]c[w+d+d']
$$
$$
	\sum_{x,y,z\in S[d,d']}
	\delta_x\delta_y\delta_z(H_{i,x}-f_x)(H_{i,y}-f_y)(H_{i,z}-f_z)
	\approx
	\sum_{w=0}^P b_i[w]b_i[w+d]b_i[w+d+d'].
$$
These convolutions can be efficiently
computed with an FFT, since under a two-dimensional discrete Fourier transform
from $(d, d')$-space to $(j, k)$-space,
$$
\sum_{w=0}^P b_i[w]b_i[w+d]b_i[w+d+d']
\leftrightarrow
B_i[j]\bar{B_i}[k]B_i[k-j],
$$
where $B_i$ is the one-dimensional discrete Fourier transform of $b$ and for
$j>0$, $B_i[-j]$ is the $j^\text{th}$ to last most element of $B_i$.
Summing over $i$ and taking the inverse discrete Fourier transform, we can
approximate the discrete Fourier transform of the numerator of $\hat{a}$. We
apply the same method to $c$ to approximate the denominator of $\hat{a}$.

@@

When only one source population is available,

>>>>

\section{Weighted Linkage Disequilibrium} As \cite{loh2013inferring} noted, we
cannot use the LD in the admixed population directly, because the allele
frequency differences in the source populations can be of either sign. As in
\cite{loh2013inferring} , we solve this problem by computing the product of the
values of the three-point linkage disequilibrium coefficient with the product of
the allele frequency differences. Using equation \ref{local_ancestry} we obtain $$
\delta_x\delta_y\delta_z
D_3(d,d')=\delta_x^2\delta_y^2\delta_z^2\mathbb{E}[\text{cov}(A_x,A_y,A_z)], $$
because the local ancestry in the admixed sample is independent of the allele
frequencies in the admixed population. For inference purposes, we estimate this
function by averaging over triples of SNPs which are separated by distances of
approximately $d$ and $d'$. The LD term is estimated from the admixed
population, while the $\delta$'s are estimated from reference populations which
are closely related to the two source populations. We notice that both this
approach, as well as the previous approaches (e.g., \cite{loh2013inferring} ),
do not take genetic drift in the source populations after the time of
admixture into account, i.e. there is an assumption of both this method and
previous methods that the allele frequencies in the ancestral source populations
can be approximated well using the allele frequencies in the extant populations.

We arrange the data from the admixed samples in an $n\times S_n $ matrix
$\mathbf{H}$, where $n$ is the number of admixed haplotypes/genotypes, and $S_n$
is the number of markers in the sample. Similarly, we arrange the data from the
two source populations into two matrices, $\mathbf{F}$ and $\mathbf{G}$,
which are of size $n_1\times S_n$ and $n_2\times S_n$, where $n_1$ and $n_2$ are
the numbers of samples from each of the source populations.
For ease of notation, we assume that the positions are given in units which make
the unit interval equal to the desired bin width.

For a given $d$ and $d'$ the SNP triples we use in the estimator for the
weighted LD are $$ S[d,d'] = \left\{x,y,z: d\leq  x-y < d+1 \text{ and } d' \leq
y-z < d'+1 \right\}. $$ Let $f_x$ be empirical allele frequency in the admixed
population. An estimator of the weighted three-point linkage disequilibrium
coefficient is then
\begin{align*}
	\hat{a}[d,d']&=
	\frac{1}{|S[d,d']|}
	\sum_{x,y,z\in S[d,d']}
		\frac{n\sum_{i=1}^n
		\hat{\delta_x}\hat{\delta_y}\hat{\delta_z}
		(H_{i,x}-f_x)(H_{i,y}-f_y)(H_{i,z}-f_z)}{(n-1)(n-2)},
\end{align*}
where
$$
	\hat{delta_x} = \sum_{i=1}^{n_1} F_{i,x} - \sum_{i=1}^{n_2} G_{i,x},
$$
and similarly for $\hat{delta_y}$ and $\hat{delta_z}$.

@@

First, rearrange $\hat{a}$ to get
\begin{align*}
	\hat{a}[d,d'] &=
	\frac{n}{(n-1)(n-2)}
	\frac{
		\sum_{i=1}^n \sum_{x,y,z\in S[d,d']}
		\hat{\delta_x} \hat{\delta_y} \hat{\delta_z}
		(H_{i,x} - f_x) (H_{i,y} - f_y) (H_{i,z} - f_z)
	}{
		\sum_{x,y,z \in S[d,d']}
		1
	},
\end{align*}
and define sequences $b_i[d]$ and $c[d]$ by binning the data and then doubling
the length by padding with $P$ zeros,
\begin{align*}
	b_i[d] &=
	\left\{
		\begin{array}{ll}
			\sum_{x: d\leq \lfloor{x}\rfloor < d+1} \hat{\delta_x}(H_{i,x}-f_x) & : 0 \leq d < P\\
			0 & : P\leq d <2P
		\end{array}
	\right.\\
	c[d] &=
	\left\{
		\begin{array}{ll}
			|\{x: d\leq \lfloor{x}\rfloor < d+1\}| & : 0\leq d < P\\
			0 & : P\leq d <2P
		\end{array}
	\right.
\end{align*}
We can approximate $|S[d,d']|$ and the $n$ sums in the numerator of
$\hat{a}[d,d']$ in terms of convolutions of these sequences:
$$
	|S[d,d']| \approx \sum_{w=0}^P c[w]c[w+d]c[w+d+d']
$$
$$
	\sum_{x,y,z\in S[d,d']}
	\hat{\delta_x} \hat{\delta_y} \hat{\delta_z}
	(H_{i,x}-f_x)(H_{i,y}-f_y)(H_{i,z}-f_z)
	\approx
	\sum_{w=0}^P b_i[w]b_i[w+d]b_i[w+d+d'].
$$
These convolutions can be efficiently
computed with an FFT, since under a two-dimensional discrete Fourier transform
from $(d, d')$-space to $(j, k)$-space,
$$
\sum_{w=0}^P b_i[w]b_i[w+d]b_i[w+d+d']
\leftrightarrow
B_i[j]\bar{B_i}[k]B_i[k-j],
$$
where $B_i$ is the one-dimensional discrete Fourier transform of $b$ and for
$j>0$, $B_i[-j]$ is the $j^\text{th}$ to last most element of $B_i$.
Summing over $i$ and taking the inverse discrete Fourier transform, we can
approximate the discrete Fourier transform of the numerator of $\hat{a}$. We
apply the same method to $c$ to approximate the denominator of $\hat{a}$.

@@

When samples only one source population is available, it is still possible to
estimate the weighted admixture linkage disequalibrium by using difference in
allele frequencies between the one source population and the admixed population
as a proxy for the difference in allele frequencies between the sampled source
population and the missing one, \cite{pickrell2012inference, loh2013inferring}.
